Supplemental data to ``\emph{k}-mer-based GWAS enhances causal variant and candidate gene discovery in soybean''

### 2023-03-01 additional_files/main_text.tex
Make minor changes to the results (mostly) so that the manuscript reads more easily

We analyzed a total of 13 traits using three GWAS approaches that differed in
the genotypes used. The first approach used SNP and indel
genotypes called by Platypus. Despite both SNPs and indels being included in
this approach, we will refer to this approach as SNP-based for simplicity. Our
SV-based workflow used various sources of data for SV discovery and a single
program (Paragraph) for genotyping SVs at population scale. Finally, the third
GWAS approach relied on the presence/absence of \textit{k}-mers as genotypes.
Because of the large number of \textit{k}-mers analyzed, we only report results
(e.g. in Manhattan plots) for those that were considered significantly
associated with traits of interest.

@@

			\csvreader[head to column names,
			tabular = llllccc,
			table head = \toprule Trait & Locus & Gene & Causal variant & Platypus\tnote{a} & Paragraph\tnote{b} & \textit{k}-mers \\ \midrule,
			table foot = \bottomrule]%
			{../tables/loci_table.csv}%
			{}%
			{\Trait & \Locus & \Gene & \Causal & \platypus & \paragraph & \kmers}

			\begin{tablenotes}
			\item[a] Platypus genotypes SNPs and indels
			\item[b] Paragraph genotypes SVs
			\item[c] Locus not discussed in main text. See Supplemental data for detailed results.
			\item[]- The approach detected no signal overlapping the gene.
			\item[]* The signal found at this locus overlaps the gene.
			\item[]** The region defined by the top 5\% (Paragraph) or 1\% (Platypus and \textit{k}-mers) variants or \textit{k}-mers overlaps the gene.
			\item[]*** The most significantly associated marker or \textit{k}-mer is the causal variant.
			\end{tablenotes}

@@

\cite{zabala2007} identified the flavonoid 3’5’-hydroxylase (F3’5’H) gene
Glyma.13g\-072100 as associated with the \textit{W1} locus for flower color. In
Williams82 and other accessions with white flowers, a 65-bp insertion into the
third exon of the gene results in a premature stop codon that renders the
F3’5’H enzyme non-functional and prevents pigmentation of the flower. All
GWAS approaches detected a signal overlapping the gene at this locus (Figures
\ref{flower_color_manhattan}, \ref{flower_color_W1_signal}). The most
significant SV and \textit{k}-mers corresponded to the known causal variant
(Figures \ref{flower-color-main-figure}a-b and \ref{flower_color_W1_gene}).
In particular, the assembly of the reads containing the most significant \textit{k}-mers
and the alignment of the resulting haplotypes showed that these \textit{k}-mers
directly tagged the causal variant (Figure \ref{flower-color-main-figure}c-e).
As the catalogue of variants analyzed by Platypus contains only SNPs and small
indels, variants located in the gene proved to be among the most closely
associated, but the causal variant could not be captured given the large size
of this variant.

@@

% Pubescence color - T locus
\cite{zabala2003} identified Glyma.06g202300 as the gene associated with the
\textit{T} locus for pubescence color and found a 1-bp deletion as
a putative causal variant at this locus. For all three GWAS approaches, we
found a signal overlapping this gene (Figure
\ref{pubescence_color_all_manhattan}). In
particular, the regions defined by the top 1\% most associated variants identified by the
Platypus and \textit{k}-mer approaches included Glyma.06g202300 (Figure
\ref{pubescence_color_all_T_signal}). While the most significant
\textit{k}-mers did occur within the sequence of this gene, they did not
correspond to the causal variant (Figure \ref{pubescence_color_all_T_gene}).
Still, the fourth and fifth most strongly associated \textit{k}-mers
corresponded to the causal variant (Figures \ref{pubescence_color_all_T_gene},
\ref{pubescence_color_all_T_kmers}). We
observed that not all accessions with gray pubescence harbored the deleted
nucleotide (Figure \ref{pubescence_color_all_T_kmers}c), which could indicate either that other causal variants are resulting in
a non-functional F3’H in this population, or that the causal
variant lies elsewhere. As for Platypus, the most significant
variant within the sequence of Glyma.06g202300 was the causal
variant, however it was not the most significantly associated variant overall
(Figures \ref{pubescence_color_all_T_signal},
\ref{pubescence_color_all_T_gene}). Paragraph did not detect the causal
variant at this locus given that this approach does not target small indels.

@@

% Pubescence color - Td locus
We conducted a separate GWAS analysis contrasting only accessions with tawny and light
tawny pubescence because the \textit{Td} locus is known to control
these differences in color. Accordingly, we
detected signals in this region using all three approaches, but only Platypus
and the \textit{k}-mers identified a signal that overlapped Glyma.03g258700
(Figures \ref{pubescence-color-main-figure}a-b,
\ref{pubescence_color_nogray_manhattan}, \ref{pubescence_color_nogray_Td_signal}).
While the most significant \textit{k}-mer identified the causal variant documented
at this locus \citep{yan2020}, it was the fourth most significantly associated variant
for the Platypus genotype calls (Figures
\ref{pubescence-color-main-figure}b-d, \ref{pubescence_color_nogray_Td_gene}).
Interestingly, some of the accessions bearing the haplotype associated with tawny pubescence
displayed light tawny pubescence, suggesting that there may be more than one causal
variant at this locus (Figure \ref{pubescence-color-main-figure}e).

@@

\begin{figure}
	\contcaption{Results of SNP/indel- and \textit{k}-mer-based GWAS at the
	\textit{Td} locus for pubescence color. (a) Genome-wide Manhattan plots.
	Horizontal dotted lines indicate the significance
	threshold whereas vertical dotted lines indicate the location of the
	\textit{Td} locus. (b) Zoomed-in Manhattan plots at the \textit{Td} locus.
	The vertical dotted lines indicate the location of the
	Glyma.03g258700 gene associated with this locus. The shaded gray
	rectangles show the regions delimited by the top 1\% associations.
	In the case of SNP/indels, blue points denote markers used in the
	original analysis, whereas red points denote points that had originally
	been pruned but whose \emph{p}-values were computed after signal
	discovery.
	(c) Gene model of Glyma.03g258700. Rectangles represent exons, colored
	rectangles represent coding sequences, and the arrow indicates the
	direction of transcription. The red rectangle highlights the region
	shown in panel (d). (d) Pairwise alignment of the two haplotype
	sequences in the population at the \textit{Td} locus. Nucleotides are
	colored according to the
	\textit{p}-value of the most significant \textit{k}-mer overlapping
	them. The vertical dotted line between haplotypes indicates the location of
	the causal SNP at this locus. (e) Contingency table of
	the phenotypes and haplotypes found in the population at the
	\textit{Td} locus.}
\end{figure}

@@

% Seed coat color - I locus
The causal variant at the \textit{I} locus for seed coat color is a complex SV
that takes the form of a 10.91-kb inverted duplication of three chalcone
synthase (CHS) genes \citep[CHS1, CHS3 and CHS4;][]{tuteja2008}. The presence of
this inverted duplication results in the silencing of all CHS genes in the seed
coat specifically and thus in the absence of seed coat pigmentation
\citep{tuteja2009}. While we did identify signals overlapping this causal variant
using all three GWAS approaches (Figures \ref{seed_coat_color_all_manhattan},
\ref{seed_coat_color_all_I_signal}), none of the approaches identified
the causal variant. This failure to detect the causal variant is likely due
to its complexity, which makes it difficult to detect and genotype it. While it
is likely that at least some of the significantly associated \textit{k}-mers
actually are derived from this causal variant, the lack of systematic methods for
linking \textit{k}-mers to sequence variation hindered our ability to do so.

@@

% Seed coat color - G locus
We conducted a second GWAS analysis contrasting accessions with yellow and
green seed coats in order to target the \textit{G} locus. Accordingly, this
analysis found strong signals overlapping the \emph{G} locus using all three
approaches (Figure \ref{seed_coat_color_greenyellow_manhattan}).
\cite{wang2018} identified Glyma.01g198500 as the gene associated with
this locus and found an A>G SNP affecting transcript splicing as the causal
variant. Consistently with their results, we identified the two
most significant \emph{k}-mers and the second most significant marker from the
Platypus analysis as corresponding to that causal variant (Figures
\ref{seed-coat-color-main-figure},
\ref{seed_coat_color_greenyellow_G_signal},
\ref{seed_coat_color_greenyellow_G_gene}). Unsurprisingly, Paragraph
did not identify the causal variant given its type.

@@

\begin{figure}
	\contcaption{Results of SNP/indel- and \textit{k}-mer-based GWAS at the
	\textit{G} locus for seed coat color. (a) Genome-wide Manhattan plots.
	Horizontal dotted lines indicate the significance
	threshold whereas vertical dotted lines indicate the location of the
	\textit{G} locus. (b) Zoomed-in Manhattan plots at the \textit{G} locus.
	The vertical dotted lines indicate the location of the
	Glyma.01g198500 gene associated with this locus. The shaded gray
	rectangles show the regions delimited by the top 1\% associations.
	In the case of SNP/indels, blue points denote markers used in the
	original analysis, whereas red points denote points that had originally
	been pruned but whose \emph{p}-values were computed after signal
	discovery.
	(c) Gene model of Glyma.01g198500. Rectangles represent exons, colored
	rectangles represent coding sequences, and the arrows indicate the
	direction of transcription. The red rectangle highlights the region
	shown in panel (d). (d) Pairwise alignment of the two haplotype
	sequences in the population at the \textit{G} locus. Nucleotides are
	colored according to the
	\textit{p}-value of the most significant \textit{k}-mer overlapping
	them. The vertical dotted line between haplotypes indicates the location of
	the causal SNP at this locus. (e) Contingency table of
	the phenotypes and haplotypes found in the population at the
	\textit{G} locus.}
\end{figure}

@@

While all GWAS approaches found signals near the Glyma.19g194300 gene
associated with the \textit{Dt1} locus for stem termination type, the
\textit{k}-mer and Platypus analyses performed best since the region defined by the most highly associated markers
(top 1\%) included the causal gene (Figures
\ref{stem_termination_all_manhattan}, \ref{stem_termination_all_Dt1_signal}).
We did identify one of the SNPs suggested as putative causal variants for this
locus by \cite{tian2010} and \cite{liu2010}, however this SNP was only the
135\textsuperscript{th} most associated marker found by Platypus and the
12\textsuperscript{th} most associated \emph{k}-mer (Figures
\ref{stem_termination_all_Dt1_gene}, \ref{stem_termination_all_Dt1_kmers}).
Our results are consistent with previous findings in
suggesting that there is more than one causal variant at the \textit{Dt1} locus \citep{liu2010, tian2010}. Indeed, a
few cultivars bearing the allele associated with \emph{dt1} showed
indeterminate or semi-determinate phenotypes and some cultivars showed a
determinate phenotype without bearing the allele associated with \emph{dt1}
(Figure \ref{stem_termination_all_Dt1_kmers}). The existence of several causal
variants might explain why our analyses did not identify a causal variant more
clearly.

% Hilum color - T locus
\subsubsection*{Hilum color -- \textit{T} locus}
\label{sv-gwas-main-results-hilum-color-t}

The \textit{T} locus mentioned above for pubescence color is also a determining
locus for hilum color. We detected signals overlapping the corresponding gene
using all approaches (Figures \ref{hilum_color_all_manhattan},
\ref{hilum_color_all_T_signal}). The regions defined by the top 5\% (Paragraph)
and top 1\% (Platypus and \textit{k}-mers) variants/\textit{k}-mers included
the Glyma.06g202300 gene associated with this locus. Similarly to what we
observed for the analysis of the \textit{T} locus based on pubescence color, we
did identify the documented causal variant at this locus using the Platypus and
\textit{k}-mer analyses, although these variants were not the most
significantly associated overall (Figures \ref{hilum_color_all_T_gene},
\ref{hilum_color_all_T_kmers}).

% Hilum color - I locus
\subsubsection*{Hilum color -- \textit{I} locus}
\label{sv-gwas-main-results-hilum-color-i}

Similarly, the \textit{I} locus discussed above for its effect on seed coat
color also controls hilum color and we detected signals overlapping the
associated inverted duplication using all approaches (Figures
\ref{hilum_color_all_manhattan}, \ref{hilum_color_all_I_signal}). However,
as was the case for the effect of \textit{I} on seed coat color, we were
not able to link any of the variants identified to the causal variation given
its highly complex nature.

@@

We conducted a GWAS analysis that only compared accessions with black and brown
hila in an attempt to detect stronger signals at the \textit{R} locus.  We
accordingly detected a signal overlapping this locus using all approaches, but
only Platypus and \textit{k}-mers detected signals overlapping the
Glyma.09g235100 gene associated with \textit{R} (Figures
\ref{hilum_color_blackbrown_manhattan}, \ref{hilum_color_blackbrown_R_signal}).
\cite{gillman2011} identified this R2R3 MYB transcription factor as the
molecular basis for the \textit{R} locus and documented four different
loss-of-function alleles linked to brown hilum color. Although our analyses did
detect strong signals in the vicinity of this gene, the most associated markers
were not located in the body of the gene (Figure
\ref{hilum_color_blackbrown_R_signal}).  Interestingly, the most significantly
associated \emph{k}-mers detected by our analysis mapped 2 kb upstream of
another gene (Glyma.09g234900) annotated as a MYB transcription factor and
putatively involved in the regulation of anthocyanin biosynthesis. This gene as
well as Glyma.09g235100 and two others are part of a tandem array of four R2R3
MYB genes considered as candidates by \cite{gillman2011}, and of which
Glyma.09g235100 was the only one expressed in the seed coat. Given the
convincing evidence provided by \cite{gillman2011}, it is most likely that
Glyma.09g235100 is indeed the gene associated with the \emph{R} locus and that
our failure to identify the most significantly associated \emph{k}-mers as
causal variants is due to the coexistence of several variants that can cause
loss of function at the \emph{R} locus.  Still, we observed two causal variants
previously documented by \cite{gillman2011} in our dataset (Figures
\ref{hilum_color_blackbrown_R_gene}, \ref{hilum_color_blackbrown_R_kmers}).

@@

We conducted a GWAS analysis comparing accessions with normal and semi-sparse
pubes\-cence in order to detect loci associated with pubescence density. This analysis
detected signals at the \textit{Ps} locus using all three approaches (Figures
\ref{pubescence-density-main-figure}a-b, \ref{pubescence_density_manhattan}, \ref{pubescence_density_Ps_signal}
and \ref{pubescence_density_Ps_gene}). The causal
variant at this recently cloned locus is a copy number variant overlapping the
Glyma.12g187200 gene, with higher copy number resulting in reduced pubescence \citep{liu2020ps}.
We identified a region of $\sim$42 kb that varied in copy number across
the population and was associated with the trait (Figure
\ref{pubescence-density-main-figure}c), and therefore possibly corresponded to the
25.6-kb region identified as the causal CNV by \cite{liu2020ps}. We found that
reads containing the most significantly associated \textit{k}-mer mapped to either of the two ends
of the CNV (Figure \ref{pubescence-density-main-figure}b,d). This suggests that
the most associated \textit{k}-mer is indeed linked to the CNV
that is causal at this locus, while \textit{k}-mers within the
CNV region may represent differences in sequence between the copies.

The difference in size between the CNV reported by \cite{liu2020ps} and our
results may be due to the presence of some unresolved sequence (a stretch of
4,635 'N' nucleotides) over that interval in the reference sequence that we used.
Because our SV filtering process removed any SV overlapping 'N' nucleotides in
the reference, we reasoned that the CNV may have been found by our pipeline but
filtered out later on. Indeed, found that a 42.8-kb duplication corresponding
to the one described above among the genotype calls made by smoove. An association
analysis of that single variant using GAPIT found a \textit{p}-value of 1.9e-23,
which would have made it the most significant SV. In this particular case, the
failure of the SV GWAS analysis to find identify the causal variant was due
to our filtering parameters and not to limitations of the SV genotype calls per se.

@@

Next, we conducted a GWAS analysis comparing dull and shiny seed coat
phenotypes for the seed coat luster trait. Seed coat luster has been shown by
\cite{gijzen1999} to be largely caused by the deposition of a hydrophobic
protein (HPS) at the seed surface. Copy number variation of this gene at the
\textit{B} locus was later shown to explain variation in luster, although
sequence variation in the HPS sequence may also play a role \citep{gijzen2006}.
Using a sequencing depth-based analysis, we identified a 31-kb region spanning
positions 9,386,109-9,417,431 on chromosome Gm15 that exhibited variation in
copy number across accessions. This CNV overlapped four genes annotated as
containing HPS domains and therefore likely represented the causal variant
(Figures \ref{seed_coat_luster_dullshiny_B_signal},
\ref{seed_coat_luster_dullshiny_B_gene}).  We found signals overlapping this
CNV using all approaches, and the most significantly associated \textit{k}-mer
was notably located within the CNV region (Figures
\ref{seed_coat_luster_dullshiny_manhattan},
\ref{seed_coat_luster_dullshiny_B_signal}). However, we were not able to
directly link any of the variants or \textit{k}-mers identified to the causal
CNV.

\subsection*{Analysis of loci with unknown causal genes}
\label{sv-gwas-results}
In addition to analyzing loci for which the genes are already known, we were
able to suggest candidate genes at loci for which the underlying genes are not
known yet.

@@

Both the \textit{L1} and the \textit{L2} loci for pod
color have yet to be cloned. While \cite{he2015} proposed the gene
Glyma.19g101700 as likely corresponding to \textit{L1} based on fine mapping
and expression analysis, our results as well as those reported by
\cite{bandillo2017} do not support this gene as a candidate due to its location
$\sim$3 Mb away from GWAS signals. Our analyses based on
\textit{k}-mers and Platypus found signals overlapping 14 and 6
genes, respectively (Figures \ref{pod_color_all_manhattan},
\ref{pod_color_all_L1_signal}). Of these, a MATE transporter (Glyma.19g120300)
appears as a prime candidate as the most significant \emph{k}-mer was observed
within its sequence and it is highly expressed in pods according to data hosted
on SoyBase \citep{grant2010, severin2010}. Moreover, the role of some MATE
proteins in transporting flavonoids has been demonstrated \citep{chen2015}.

The signals found by Platypus and \emph{k}-mer analyses at the \textit{L2}
locus overlap 24 and 50 genes, respectively (Figures
\ref{pod_color_all_manhattan}, \ref{pod_color_all_L2_signal}). Despite this
somewhat high number, we noticed that the closest gene to the
most significant \emph{k}-mer (Glyma.03g005800) is also a MATE transporter,
similarly to our top candidate for \emph{L1}. This gene may therefore represent
an interesting candidate for this locus for future studies.

@@

For the GWAS analyses computed from Platypus genotype calls, we retrieved all
markers that had initially been pruned in a region ranging from 50 kb
downstream to 50 kb upstream of significant signals and computed their
\emph{p}-values by running GAPIT using the same parameters as noted above. The
\emph{p}-values thus computed on markers that were initially left out were not
used for initial signal discovery, but provided finer resolution when analyzing
the results of the Platypus GWAS analyses. Indeed, some markers that had not
been included in the original GWAS may be the causal variant or simply yield
more significant \emph{p}-values than those initially used.

>>>>

We analyzed a total of 13 traits in a population of 363 \textit{G. max}
accessions using three GWAS approaches that differed in the genotypes used. The
first approach used SNP and indel genotypes called by Platypus. Despite both
SNPs and indels being included in this approach, we will refer to this approach
as SNP-based for simplicity. Our SV-based workflow used various sources of data
for SV discovery and a single program (Paragraph) for genotyping SVs at
population scale. Finally, the third GWAS approach relied on the
presence/absence of \textit{k}-mers as genotypes.  Because of the large number
of \textit{k}-mers analyzed, we only report results (e.g. in Manhattan plots)
for those that were considered significantly associated with traits of
interest.

@@

			\csvreader[head to column names,
			tabular = llllccc,
			table head = \toprule Trait & Locus & Gene & Causal variant & SNPs\tnote{a} & SVs\tnote{b} & \textit{k}-mers \\ \midrule,
			table foot = \bottomrule]%
			{../tables/loci_table.csv}%
			{}%
			{\Trait & \Locus & \Gene & \Causal & \platypus & \paragraph & \kmers}

			\begin{tablenotes}
			\item[a] SNPs and indels genotyped by Platypus
			\item[b] SVs genotyped by Paragraph
			\item[c] Locus not discussed in main text. See Supplemental data for detailed results.
			\item[]- The approach detected no signal overlapping the gene.
			\item[]* The signal found at this locus overlaps the gene.
			\item[]** The region defined by the top 5\% (SVs) or 1\% (SNPs and \textit{k}-mers) variants overlaps the gene.
			\item[]*** The most significantly associated variant is the causal variant.
			\end{tablenotes}

@@

\cite{zabala2007} identified the flavonoid 3’5’-hydroxylase (F3’5’H) gene
Glyma.13g\-072100 as associated with the \textit{W1} locus for flower color. In
Williams82 and other accessions with white flowers, a 65-bp insertion into the
third exon of the gene results in a premature stop codon that renders the
F3’5’H enzyme non-functional and prevents pigmentation of the flower. All
GWAS approaches detected a signal overlapping the gene at this locus (Figures
\ref{flower_color_manhattan}, \ref{flower_color_W1_signal}). The most
significant SV and \textit{k}-mers corresponded to the known causal variant
(Figures \ref{flower-color-main-figure}a-b and \ref{flower_color_W1_gene}).
In particular, the assembly of the reads containing the most significant \textit{k}-mers
and the alignment of the resulting haplotypes showed that these \textit{k}-mers
directly tagged the causal variant (Figure \ref{flower-color-main-figure}c-e).
As the catalogue of variants analyzed by Platypus contains only SNPs and small
indels, the causal variant could not be captured by this approach. Still, SNPs
and indels located within the gene sequence were among the most significantly
associated ones.

@@

% Pubescence color - T locus
\cite{zabala2003} identified Glyma.06g202300 as the gene associated with the
\textit{T} locus for pubescence color and found a 1-bp deletion as a putative
causal variant at this locus. For all three GWAS approaches, we found a signal
overlapping this gene (Figure \ref{pubescence_color_all_manhattan}). In
particular, the regions defined by the 1\% most associated variants
identified by the SNP- and \textit{k}-mer-based approaches included
Glyma.06g202300 (Figure \ref{pubescence_color_all_T_signal}). While the most
significant \textit{k}-mers did occur within the sequence of this gene, they
did not correspond to the causal variant (Figure
\ref{pubescence_color_all_T_gene}).  Still, the fourth and fifth most strongly
associated \textit{k}-mers corresponded to the causal variant (Figures
\ref{pubescence_color_all_T_gene}, \ref{pubescence_color_all_T_kmers}). We
observed that not all accessions with gray pubescence harbored the deleted
nucleotide (Figure \ref{pubescence_color_all_T_kmers}c), which could indicate
either that other causal variants are resulting in a non-functional F3’H in
this population, or that the causal variant lies elsewhere. As for SNP-based GWAS,
the most significant variant within the sequence of Glyma.06g202300 was the
causal variant, however it was not the most significantly associated variant
overall (Figures \ref{pubescence_color_all_T_signal},
\ref{pubescence_color_all_T_gene}). SV-based GWAS did not detect the causal variant
at this locus, as expected from its size.

@@

% Pubescence color - Td locus
We conducted a separate GWAS analysis contrasting only accessions with tawny and light
tawny pubescence because the \textit{Td} locus is known to control
these differences in color. Accordingly, we
detected signals in this region using all three approaches, but only SNP-
and \textit{k}-mer-based GWAS identified a signal that overlapped Glyma.03g258700
(Figures \ref{pubescence-color-main-figure}a-b,
\ref{pubescence_color_nogray_manhattan}, \ref{pubescence_color_nogray_Td_signal}).
While the most significant \textit{k}-mer identified the causal variant documented
at this locus \citep{yan2020}, it was the fourth most significantly associated variant
for the SNP-based approach (Figures
\ref{pubescence-color-main-figure}b-d, \ref{pubescence_color_nogray_Td_gene}).
Interestingly, some of the accessions bearing the haplotype associated with tawny pubescence
displayed light tawny pubescence, suggesting that there may be more than one causal
variant at this locus (Figure \ref{pubescence-color-main-figure}e).

@@

\begin{figure}
	\contcaption{Results of SNP- and \textit{k}-mer-based GWAS at the
	\textit{Td} locus for pubescence color. (a) Genome-wide Manhattan plots.
	Horizontal dotted lines indicate the significance
	threshold whereas vertical dotted lines indicate the location of the
	\textit{Td} locus. (b) Zoomed-in Manhattan plots at the \textit{Td} locus.
	The vertical dotted lines indicate the location of the
	Glyma.03g258700 gene associated with this locus. The shaded gray
	rectangles show the regions delimited by the top 1\% associations.
	In the case of SNPs, blue points denote markers used in the
	original analysis, whereas red points denote points that had originally
	been pruned but whose \emph{p}-values were computed after signal
	discovery.
	(c) Gene model of Glyma.03g258700. Rectangles represent exons, colored
	rectangles represent coding sequences, and the arrow indicates the
	direction of transcription. The red rectangle highlights the region
	shown in panel (d). (d) Pairwise alignment of the two haplotype
	sequences in the population at the \textit{Td} locus. Nucleotides are
	colored according to the
	\textit{p}-value of the most significant \textit{k}-mer overlapping
	them. The vertical dotted line between haplotypes indicates the location of
	the causal SNP at this locus. (e) Contingency table of
	the phenotypes and haplotypes found in the population at the
	\textit{Td} locus.}
\end{figure}

@@

% Seed coat color - I locus
The causal variant at the \textit{I} locus for seed coat color is a complex SV
that takes the form of a 10.91-kb inverted duplication of three chalcone
synthase (CHS) genes \citep[CHS1, CHS3 and CHS4;][]{tuteja2008}. The presence of
this inverted duplication results in the silencing of all CHS genes in the seed
coat specifically and thus in the absence of seed coat pigmentation
\citep{tuteja2009}. While we did find signals overlapping this causal variant
using all three GWAS approaches (Figures \ref{seed_coat_color_all_manhattan},
\ref{seed_coat_color_all_I_signal}), none of the approaches identified
it. This failure to detect the causal variant is likely due
to its complexity, which makes it difficult to detect and genotype it. While it
is likely that at least some of the significantly associated \textit{k}-mers
actually are derived from this causal variant, the lack of systematic methods for
linking \textit{k}-mers to sequence variation hindered our ability to do so.

@@

% Seed coat color - G locus
We conducted a second GWAS analysis contrasting accessions with yellow and
green seed coats in order to target the \textit{G} locus. Accordingly, this
analysis found strong signals overlapping the \emph{G} locus using all three
approaches (Figure \ref{seed_coat_color_greenyellow_manhattan}).
\cite{wang2018} identified Glyma.01g198500 as the gene associated with
this locus and found an A>G SNP affecting transcript splicing as the causal
variant. Consistently with their results, we identified the two
most significant \emph{k}-mers and the second most significant marker from the
SNP-based analysis as corresponding to that causal variant (Figures
\ref{seed-coat-color-main-figure},
\ref{seed_coat_color_greenyellow_G_signal},
\ref{seed_coat_color_greenyellow_G_gene}). Unsurprisingly, SV-based GWAS
did not identify the causal variant given its type.

@@

\begin{figure}
	\contcaption{Results of SNP- and \textit{k}-mer-based GWAS at the
	\textit{G} locus for seed coat color. (a) Genome-wide Manhattan plots.
	Horizontal dotted lines indicate the significance
	threshold whereas vertical dotted lines indicate the location of the
	\textit{G} locus. (b) Zoomed-in Manhattan plots at the \textit{G} locus.
	The vertical dotted lines indicate the location of the
	Glyma.01g198500 gene associated with this locus. The shaded gray
	rectangles show the regions delimited by the top 1\% associations.
	In the case of SNPs, blue points denote markers used in the
	original analysis, whereas red points denote points that had originally
	been pruned but whose \emph{p}-values were computed after signal
	discovery.
	(c) Gene model of Glyma.01g198500. Rectangles represent exons, colored
	rectangles represent coding sequences, and the arrows indicate the
	direction of transcription. The red rectangle highlights the region
	shown in panel (d). (d) Pairwise alignment of the two haplotype
	sequences in the population at the \textit{G} locus. Nucleotides are
	colored according to the
	\textit{p}-value of the most significant \textit{k}-mer overlapping
	them. The vertical dotted line between haplotypes indicates the location of
	the causal SNP at this locus. (e) Contingency table of
	the phenotypes and haplotypes found in the population at the
	\textit{G} locus.}
\end{figure}

@@

While all GWAS approaches found signals near the Glyma.19g194300 gene
associated with the \textit{Dt1} locus for stem termination type, the SNP- and
\textit{k}-mer-based analyses performed best since the region defined by their
most highly associated markers (top 1\%) included the causal gene (Figures
\ref{stem_termination_all_manhattan}, \ref{stem_termination_all_Dt1_signal}).
We did identify one of the SNPs suggested as putative causal variants for this
locus by \cite{tian2010} and \cite{liu2010}, however this SNP was only the
135\textsuperscript{th} most associated marker found by Platypus and the
12\textsuperscript{th} most associated \emph{k}-mer (Figures
\ref{stem_termination_all_Dt1_gene}, \ref{stem_termination_all_Dt1_kmers}).
Our results are consistent with previous findings in suggesting that there is
more than one causal variant at the \textit{Dt1} locus \citep{liu2010,
tian2010}. Indeed, a few cultivars bearing the allele associated with
\emph{dt1} showed indeterminate or semi-determinate phenotypes and some
cultivars showed a determinate phenotype without bearing the allele associated
with \emph{dt1} (Figure \ref{stem_termination_all_Dt1_kmers}). The existence of
several causal variants might explain why our analyses did not identify a
causal variant more clearly.

@@

We conducted a GWAS analysis that only compared accessions with black and brown
hila in an attempt to detect stronger signals at the \textit{R} locus.  We
accordingly detected a signal overlapping this locus using all approaches, but
only SNP- and \textit{k}-mer-based GWAS detected signals overlapping the
Glyma.09g235100 gene associated with \textit{R} (Figures
\ref{hilum_color_blackbrown_manhattan}, \ref{hilum_color_blackbrown_R_signal}).
\cite{gillman2011} identified this R2R3 MYB transcription factor as the
molecular basis for the \textit{R} locus and documented four different
loss-of-function alleles linked to brown hilum color. Although our analyses did
detect strong signals in the vicinity of this gene, the most associated markers
were not located in the body of the gene (Figure
\ref{hilum_color_blackbrown_R_signal}).  Interestingly, the most significantly
associated \emph{k}-mers detected by our analysis mapped 2 kb upstream of
another gene (Glyma.09g234900) annotated as a MYB transcription factor and
putatively involved in the regulation of anthocyanin biosynthesis. This gene as
well as Glyma.09g235100 and two others are part of a tandem array of four R2R3
MYB genes considered as candidates by \cite{gillman2011}, and of which
Glyma.09g235100 was the only one expressed in the seed coat. Given the
convincing evidence provided by \cite{gillman2011}, it is most likely that
Glyma.09g235100 is indeed the gene associated with the \emph{R} locus and that
our failure to identify the most significantly associated \emph{k}-mers as
causal variants is due to the coexistence of several variants that can cause
loss of function at the \emph{R} locus.  Still, we observed two causal variants
previously documented by \cite{gillman2011} in our dataset (Figures
\ref{hilum_color_blackbrown_R_gene}, \ref{hilum_color_blackbrown_R_kmers}).

@@

We conducted a GWAS analysis comparing accessions with normal and semi-sparse
pubes\-cence in order to detect loci associated with pubescence density. This analysis
detected signals at the \textit{Ps} locus using all three approaches (Figures
\ref{pubescence-density-main-figure}a-b, \ref{pubescence_density_manhattan}, \ref{pubescence_density_Ps_signal}
and \ref{pubescence_density_Ps_gene}). The causal
variant at this recently cloned locus is a copy number variant (CNV) overlapping the
Glyma.12g187200 gene, with higher copy number resulting in reduced pubescence \citep{liu2020ps}.
We identified a region of $\sim$42 kb that varied in copy number across
the population and was associated with the trait (Figure
\ref{pubescence-density-main-figure}c), and therefore possibly corresponded to the
25.6-kb region identified as the causal CNV by \cite{liu2020ps}. We found that
reads containing the most significantly associated \textit{k}-mer mapped to either of the two ends
of the CNV (Figure \ref{pubescence-density-main-figure}b,d). This suggests that
the most associated \textit{k}-mer is indeed linked to the CNV
that is causal at this locus, while \textit{k}-mers within the
CNV region may represent differences in sequence between the copies.

The difference in size between the CNV reported by \cite{liu2020ps} and our
results may be due to the presence of some unresolved sequence (a stretch of
4,635 'N' nucleotides) over that interval in the reference sequence that we used.
Because our SV filtering process removed any SV overlapping 'N' nucleotides in
the reference, we reasoned that the CNV may have been found by our pipeline but
filtered out later on. Indeed, found a 42.8-kb duplication corresponding
to the one described above among the genotype calls made by smoove. An association
analysis of that single variant using GAPIT found a \textit{p}-value of 1.9e-23,
which would have made it the most significant SV. In this particular case, the
failure of the SV-based GWAS analysis to find identify the causal variant was due
to our filtering parameters and not to limitations of the SV genotype calls per se.

@@

We conducted a GWAS analysis comparing dull and shiny seed coat phenotypes for
the seed coat luster trait. Seed coat luster has been shown by
\cite{gijzen1999} to be largely caused by the deposition of a hydrophobic
protein (HPS) at the seed surface. Copy number variation of this gene at the
\textit{B} locus was later shown to explain variation in luster, although
sequence variation in the HPS sequence may also play a role \citep{gijzen2006}.
Using a sequencing depth-based analysis, we identified a 31-kb region spanning
positions 9,386,109-9,417,431 on chromosome Gm15 that exhibited variation in
copy number across accessions. This CNV overlapped four genes annotated as
containing HPS domains and therefore likely represented the causal variant
(Figures \ref{seed_coat_luster_dullshiny_B_signal},
\ref{seed_coat_luster_dullshiny_B_gene}).  We found signals overlapping this
CNV using all approaches, and the most significantly associated \textit{k}-mer
was notably located within the CNV region (Figures
\ref{seed_coat_luster_dullshiny_manhattan},
\ref{seed_coat_luster_dullshiny_B_signal}). However, we were not able to
directly link any of the variants or \textit{k}-mers identified to the causal
CNV.

\subsection*{Analysis of loci with unknown causal genes}
\label{sv-gwas-results}

In addition to analyzing loci for which the genes are already known, we were
able to suggest candidate genes at loci for which the underlying genes are not
known yet.

@@

Both the \textit{L1} and the \textit{L2} loci for pod color have yet to be
cloned. While \cite{he2015} proposed the gene Glyma.19g101700 as likely
corresponding to \textit{L1} based on fine mapping and expression analysis, our
results as well as those reported by \cite{bandillo2017} do not support this
gene as a candidate due to its location $\sim$3 Mb away from GWAS signals. Our
analyses based on \textit{k}-mers and SNPs found signals overlapping 14 and
6 genes, respectively (Figures \ref{pod_color_all_manhattan},
\ref{pod_color_all_L1_signal}). Of these, a MATE transporter (Glyma.19g120300)
appears as a prime candidate as the most significant \emph{k}-mer was observed
within its sequence and it is highly expressed in pods according to data hosted
on SoyBase \citep{grant2010, severin2010}. Moreover, the role of some MATE
proteins in transporting flavonoids has been demonstrated \citep{chen2015}.

The signals found by SNP- and \emph{k}-mer-based analyses at the \textit{L2}
locus overlap 24 and 50 genes, respectively (Figures
\ref{pod_color_all_manhattan}, \ref{pod_color_all_L2_signal}). Despite this
somewhat high number, we noticed that the closest gene to the most significant
\emph{k}-mer (Glyma.03g005800) is also a MATE transporter, similarly to our top
candidate for \emph{L1}. This gene may therefore represent an interesting
candidate for this locus for future studies.

@@

For the GWAS analyses computed from Platypus genotype calls, we retrieved all
markers that had initially been pruned in a region ranging from 50 kb
downstream to 50 kb upstream of significant signals and computed their
\emph{p}-values by running GAPIT using the same parameters as noted above. The
\emph{p}-values thus computed on markers that were initially left out were not
used for initial signal discovery, but provided finer resolution when analyzing
the results of the SNP-based GWAS analyses. Indeed, some markers that had not
been included in the original GWAS may be the causal variant or simply yield
more significant \emph{p}-values than those initially used.
